DATA 607 Project 1 - Chess Scores

### 2018-09-14 DATA607/Project_1.Rmd
Additional work on Project 1

## Setting up My Environment

@@

Now to explain my apporach.  I will loop through the character vector.  I will check if the line is made up of dashes.  After reading two dashed lines I will start extracting data.  There are two lines that have useful data.  The first has the player's id, name and match data.  The second line has the player's state and starting score.

@@

```{r the_loop}
# Loop through line by line
for (line in lines){
  # Look for the dash line
  if (line == "-----------------------------------------------------------------------------------------"){
    # Increment the dash line count
    dash_lines_found <- dash_lines_found + 1
    # Check if we need to reset the data line counter
    if (extract_data){
      data_line_count <- 0
    }
    # Check if we need to extract data
    if (dash_lines_found == 2){
      extract_data <- TRUE
    }
  } else if (extract_data) {
    # Increase the data line count by one
    data_line_count <- data_line_count + 1
    # Split the data by the pipe character
    split_line <- unlist(strsplit(line, "\\|"))
    # Extract the data
    if (data_line_count == 1){
      # This is the first line so extract the player's id
      player_id <- str_extract(split_line[1], "[[:digit:]]+")
      # Extract name, split name into parts, capitalize the first letter and
      # put it back together
      player_name <- str_extract(split_line[2], "[[:alpha:]]+ [[:alpha:]]+")
      player_name <- unlist(strsplit(player_name, " "))
      player_name <- capitalize_first_letter(player_name)
      player_name <- paste(player_name, collapse = " ")
      # Loop through the match results
      for (game in split_line[4:10]){
        # Extract if it is a win, loss, draw, etc.
        result <- str_extract(game, "[[:alpha:]]")
        # Extract the opponent's id
        opponent_id <- str_extract(game, "[[:digit:]]+")
        # Add the data to the match vector
        matches <- rbind(matches, c(player_id, opponent_id, result))
      }
    } else {
      # This is the second line so Get the player's rating
      player_rating <- unlist(str_extract_all(split_line[2], "[[:digit:]]+"))
      player_rating <- as.numeric(player_rating[2])
      # Get the player's state
      player_state <- str_extract(split_line[1], "[[:alpha:]]+")
      # Pull all the extracted player data together into a list
      player_data <- list("player_id" = player_id,
                          "players_name" = player_name,
                          "players_state" = player_state,
                          "players_pre_rating" = player_rating)
      # Turn that list into a dataframe with one row
      player_row <- data.frame(bind_rows(player_data))
      # Add it to the players dataframe
      if (!exists("player_df")){
        player_df <- player_row
      } else {
        player_df <- rbind(player_df, player_row)
      }
    }
  }
}
```

>>>>

## Setting Up My Environment

@@

I will loop through the character vector.  I will check if the line is made up of dashes.  After reading two dashed lines I will start extracting data.  There are two lines that have useful data.  The first has the player's id, name and match data.  The second line has the player's state and starting score.

@@

```{r the_loop}
# Loop through line by line
for (line in lines){
  # Look for the dash line
  if (line == "-----------------------------------------------------------------------------------------"){
    # Increment the dash line count
    dash_lines_found <- dash_lines_found + 1
    # Check if we need to reset the data line counter
    if (extract_data){
      data_line_count <- 0
    }
    # Check if we need to extract data
    if (dash_lines_found == 2){
      extract_data <- TRUE
    }
  } else if (extract_data) {
    # Increase the data line count by one
    data_line_count <- data_line_count + 1
    # Split the data by the pipe character
    split_line <- unlist(strsplit(line, "\\|"))
    # Extract the data
    if (data_line_count == 1){
      # This is the first line so extract the player's id
      player_id <- str_extract(split_line[1], "[[:digit:]]+")
      # Extract name, split name into parts, capitalize the first letter and
      # put it back together
      player_name <- str_extract_all(split_line[2], "[[:alpha:]]+")
      player_name <- unlist(player_name)
      player_name <- capitalize_first_letter(player_name)
      player_name <- paste(player_name, collapse = " ")
      # Extract total number of points
      player_total_points <- as.numeric(trimws(split_line[3]))
      # Loop through the match results
      for (game in split_line[4:10]){
        # Extract if it is a win, loss, draw, etc.
        result <- str_extract(game, "[[:alpha:]]")
        # Extract the opponent's id
        opponent_id <- str_extract(game, "[[:digit:]]+")
        # Add the data to the match vector
        matches <- rbind(matches, c(player_id, opponent_id, result))
      }
    } else {
      # This is the second line so Get the player's rating
      player_rating <- unlist(str_extract_all(split_line[2], "[[:digit:]]+"))
      player_rating <- as.numeric(player_rating[2])
      # Get the player's state
      player_state <- str_extract(split_line[1], "[[:alpha:]]+")
      # Pull all the extracted player data together into a list
      player_data <- list("player_id" = player_id,
                          "players_name" = player_name,
                          "players_state" = player_state,
                          "player_total_points" = player_total_points,
                          "players_pre_rating" = player_rating)
      # Turn that list into a dataframe with one row
      player_row <- data.frame(bind_rows(player_data))
      # Add it to the players dataframe
      if (!exists("player_df")){
        player_df <- player_row
      } else {
        player_df <- rbind(player_df, player_row)
      }
    }
  }
}
```
